---
title: "Assignment 3"
author: "Daniel Glen Cox"
date: "10 April 2017"
output: 
  pdf_document:
    fig_caption: true
font_size: 12pt
---

```{r echo=FALSE,message=FALSE,include=FALSE}
library(mgcv)
library(INLA)
library(Pmisc)
INLA:::inla.dynload.workaround()

# Non-parametrics
cUrl=paste("http://scrippsco2.ucsd.edu/sites/default/files","/data/flask_co2_and_isotopic/daily_co2/fldav_spo.csv",sep="")
cFile=basename(cUrl)
if(!file.exists(cFile))download.file(cUrl, cFile)
co2s=read.table(cFile,header=FALSE,sep=",",skip=69,stringsAsFactors=FALSE)
co2s[co2s[,6]>0,7]=NA
co2s=data.frame(date=strptime(co2s[,1],format="%Y-%m-%d",tz="UTC"),co2=co2s[,7])
timeOrigin=ISOdate(1980,1,1,0,0,0,tz="UTC")
co2s$days=as.numeric(difftime(co2s$date,timeOrigin,units="days"))
co2s$cos12=cos(2*pi*co2s$days/365.25)
co2s$sin12=sin(2*pi*co2s$days/365.25)
co2s$cos6=cos(2*2*pi*co2s$days/365.25)
co2s$sin6=sin(2*2*pi*co2s$days/365.25)

# the models

cgam1=gam(co2~s(days) + cos12 + sin12 + cos6 + sin6,data=co2s)

# predictions

eps=1
newX=co2s
newXfwd=newX+eps
newXbwd=newX-eps
pred1=predict(cgam1,newXbwd,type='lpmatrix')
pred2=predict(cgam1,newX,type='lpmatrix')
pred3=predict(cgam1,newXfwd,type='lpmatrix')

fd21=(pred2-pred1)/eps
fd21=fd21%*%coef(cgam1)
sdiff=(pred3+pred1-2*pred2)/(eps^2)
sdiff=sdiff%*%coef(cgam1)

#og code
newXpred=data.frame(date=seq(ISOdate(1990,1,1,0,0,0,tz ="UTC"),by="1 days",length.out=365*30))
newXpred$days=as.numeric(difftime(newXpred$date, timeOrigin,units="days"))
newXpred$cos12=cos(2*pi*newXpred$days/365.25)
newXpred$sin12=sin(2*pi*newXpred$days/365.25)
newXpred$cos6=cos(2*2*pi*newXpred$days/365.25)
newXpred$sin6=sin(2*2*pi*newXpred$days/365.25)
coPred=predict(cgam1,newXpred,se.fit=TRUE)
coPred=data.frame(est=coPred$fit,lower=coPred$fit-2*coPred$se.fit,upper=coPred$fit+2*coPred$se.fit)
newXannual=newXpred[1:365, ]
newXannual$days=0
```

# Non-parametrics

## Introduction

Carbon dioxide has long been known to be a greenhouse gas, 
trapping heat within the atmosphere. More contentious, however, 
is the assertion that carbon dioxide concentrations in the 
atmosphere are increasing as a result of human actions. In an 
effort to explore this issue in detail, this study sought to 
answer the following questions regarding carbon dioxide levels 
in Antarctica:

* Even though carbon dioxide levels are still increasing, has 
the increase slowed somewhat recently?

* Did global economic recessions around 1980-1982 and 1990 
cause a slowing in the increase of carbon dioxide levels?

* Are carbon levels higher in October than in March?

* Are carbon dioxide levels likely to exceed 400 parts per 
gallon by 2020?

To explore these hypotheses, this study made use of data from 
the Scripps CO2 Program, generously provided to the public by 
the Scripps Institute of Oceanography at the University of 
California in San Diego.

## Methods

This study employed a generalized additive model, defined as follows:
$$
\begin{aligned}
Y_{i} &\sim N(\mu_{i}, \tau^{2}) \\
\mu_{i} &= X_{i}\beta + f(W_{i})
\end{aligned}
$$
Here $W_{i}$ is a covariate for days, which have been turned into a day count 
from the original date variable, and $f$ is a smooth function of $W_{i}$. 
The covariates $X_{i}$ consist of variables which effectively model annual and 
bi-annual cycles.

The data itself in its original form consists of dates with corresponding 
carbon dioxide measurements. The data set was then appended with a countdown 
of days to present time and columns containing the results of sine and 
cosine functions with annual and bi-annual cycles evaluated on the day counts.

## Results

The results for the model fit itself can be found below in table 1.

```{r echo=FALSE}
knitr::kable(summary(cgam1)$p.table,digits=3,
             caption='Generalized additive model of carbon concentration')
```

```{r echo=FALSE}
# Plot of prediction to 2020
plot(newXpred$date, coPred$est,type="l",main="Prediction of Concentration to 2020",
     xlab="Year",ylab="Concentration")
matlines(as.numeric(newXpred$date), coPred[,c("lower","upper","est")],lty =1,col=c("yellow","yellow","black"))
```

```{r echo=FALSE}
# Plot of annual trend
plot(newXannual$date,predict(cgam1, newXannual),main="Annual Trend",
     xlab="Months",ylab="Concentration")
```

```{r echo=FALSE}
plot(newX$date,fd21,main='Approximation of First Derivative',
xlab='Year',ylab='First Difference')
```

```{r echo=FALSE}
plot(newX$date,sdiff,main='Approximation of Second Derivative',
xlab='Year',ylab='Second Difference')
lines(newX$date,rep(0,dim(newX)[1]),col='red',lwd='2')
```

## Discussion



# Moss in Galicia: Redux



# Application

## Introduction

The use of tobacco products is both a major health concern, as well as 
a social one. One of the more complex aspects of tobacco consumption 
that many researchers seek to understand how an individual acquires a 
tobacco habit in the first place. To help explore one of the many 
facets to that issue, this study sought to examine the age at which 
children first try smoking cigarettes, and how that age might vary 
amongst different regions and other demographics. In particular, this 
study posed the following questions:

To this end, this study made use of the 2014 American National Youth 
Tobacco Survey. Details regarding this dataset are given in the methods 
section.

## Methods

The data used to carry out this study comes from the 2014 American National 
Youth Tobacco Survey. It was carried out with children in grades 6 to 12 and 
comprises 81 questions on tobacco-related beliefs, attitudes, and behaviours, 
as well as questions on exposure to pro- and anti-tobacco influences. The 
survey had 22,007 respondents.

## Results



## Discussion

